---
title: "Analysis of Film and Music Sales for Media Distributors, Inc."
author: "Name: Xujia Qin"
date: "Date: SummerI 2025"
output:
  html_document:
    toc: true
    df_print: paged
  html_notebook:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
      toc_depth: 2
    number_sections: true
    theme: paper
    highlight: tango
subtitle: Prepared by Oakland Partners
---


```{r echo = FALSE, message=FALSE, warning=FALSE}
library(DBI)
library(RMySQL) 
library(ggplot2)
library(kableExtra)
library(knitr)

# Connect to the database
db_user <- 'avnadmin'
db_password <- 'AVNS_myNMlu7wBu69NOuFkU1'
db_name <- 'defaultdb'
db_host <- 'mysql-1c2ba4a8-northeastern-cee5.k.aivencloud.com'
db_port <- 19461

con <- dbConnect(MySQL(), user = db_user, password = db_password,
                 dbname = db_name, host = db_host, port = db_port)
```

### 1. Background

Media Distributors, Inc. is a Wichita, KY based distributor and seller of films and music for commercial purposes. For the past few years, it has managed its film and music sales separately using different applications. This division is historical as Media Distributors started distributing films and then acquired SoundMania about two years ago. As the two distribution channels were different, CTO Alvin Coelho made the decision not to integrate the SoundMania’s information systems and database with those of Media Distributors. This has not been a problem until now, however Media Distributors intends to make itself available for acquisition. To that end, an integrated view of the business, particularly sales, is needed.

This report provides key sales, revenue, and customer metrics to showcase Media Distributors business. The analysis provided is based on data from a custom-built datamart containing data extracted from two operational databases.


### 2. Key Business Metrics

This sections provides key revenue and customer information segmented by time, country, and business unit. Revenue numbers are in US$.

#### 2.1 Sales Revenue
```{r echo = FALSE, message=FALSE, warning=FALSE}
# Get total revenue per year
yearly_revenue <- dbGetQuery(con, "
  SELECT YEAR(date) AS year, ROUND(SUM(revenue), 0) AS total_revenue
  FROM fact_sales
  GROUP BY year
  ORDER BY total_revenue DESC;
")

# Get the top revenue year and value
top_year <- yearly_revenue$year[1]
top_revenue <- formatC(yearly_revenue$total_revenue[1], format = "d", big.mark = ",")

# Get revenue by country/year
country_year_revenue <- dbGetQuery(con, "
  SELECT country, YEAR(date) AS year, ROUND(SUM(revenue), 0) AS revenue
  FROM fact_sales
  GROUP BY country, year
  ORDER BY revenue DESC;
")

# Most revenue by country/year
top_country_row <- country_year_revenue[1, ]
top_country <- top_country_row$country
top_country_year <- top_country_row$year
top_country_revenue <- formatC(top_country_row$revenue, format = "d", big.mark = ",")

# Second place country
second_country_row <- country_year_revenue[2, ]
second_country <- second_country_row$country
second_country_revenue <- formatC(second_country_row$revenue, format = "d", big.mark = ",")

# Extract two most recent years from fact_sales
most_recent_years <- dbGetQuery(con, "
  SELECT DISTINCT YEAR(date) AS year
  FROM fact_sales
  ORDER BY year DESC
  LIMIT 2;
")

# Extract and assign years
year1 <- most_recent_years$year[1]
year2 <- most_recent_years$year[2]


```

The year with the most sales was `r top_year` with a total revenue across both business units of $`r top_revenue`.
The country with the most sales was `r top_country` with total sales across both business units in `r top_country_year` of $`r top_country_revenue`. It was followed by `r second_country` with $`r second_country_revenue`.

The table below shows sales revenue by country and ordered by total sales for the two most recent years, `r year1` and `r year2`.The table is restricted to the top five countries with the most sales. The column ‘Total’ in the table represents the total for all years for which there is data and not just the past two years.

```{r echo = FALSE, message=FALSE, warning=FALSE}
# Get the two most recent years
most_recent_years <- dbGetQuery(con, "
  SELECT DISTINCT YEAR(date) AS year
  FROM fact_sales
  ORDER BY year DESC
  LIMIT 2;
")
year1 <- most_recent_years$year[1]
year2 <- most_recent_years$year[2]

# Query revenue data for only those two years
revenue_by_country <- dbGetQuery(con, sprintf("
  SELECT 
    country, 
    YEAR(date) AS year, 
    ROUND(SUM(revenue), 0) AS revenue
  FROM fact_sales
  WHERE YEAR(date) IN (%d, %d)
  GROUP BY country, year
  ORDER BY year DESC;
", year1, year2))

# Reshape to wide format
revenue_wide <- reshape(revenue_by_country,
                        idvar = "country",
                        timevar = "year",
                        direction = "wide")

# Replace NA with 0s
revenue_wide[is.na(revenue_wide)] <- 0

# Find revenue columns
revenue_cols <- grep("^revenue\\.", colnames(revenue_wide))

# Add total column
revenue_wide$Total <- rowSums(revenue_wide[, revenue_cols])

# Rename columns like revenue.2022 → 2022
colnames(revenue_wide)[revenue_cols] <- sub("revenue\\.", "", colnames(revenue_wide)[revenue_cols])

# Get top 5 countries by total revenue
top5_countries <- head(revenue_wide[order(-revenue_wide$Total), ], 5)

# Final column names for display
col_headers <- c("Country", as.character(c(year2, year1)), "Total (All Years)")

# Display the table
kbl(top5_countries,
    caption = "Total Revenue For Top Five Countries For Most Recent Two Years",
    col.names = col_headers) %>%
  kable_styling("striped", full_width = FALSE, position = "center")


```


```{r echo = FALSE, message=FALSE, warning=FALSE}
# Get min and max year from fact_sales
year_range <- dbGetQuery(con, "
  SELECT MIN(YEAR(date)) AS start_year, MAX(YEAR(date)) AS end_year
  FROM fact_sales;
")
start_year <- year_range$start_year
end_year <- year_range$end_year
```
The table below shows the revenue broken down by quarter for the top five countries.  It shows the total revenue for each quarter across all business units and years.  So, for example, the column “Q1” is the total sales for music and film for all years for which there is data, which is from `r start_year` to `r end_year`.

```{r echo = FALSE, message=FALSE, warning=FALSE}
# Aggregate revenue by country and quarter across all years
quarterly_rev <- dbGetQuery(con, "
  SELECT 
    country,
    QUARTER(date) AS quarter,
    ROUND(SUM(revenue), 0) AS revenue
  FROM fact_sales
  GROUP BY country, quarter
")

# Reshape to wide format: one row per country, columns Q1 to Q4
quarterly_wide <- reshape(quarterly_rev,
                          idvar = "country",
                          timevar = "quarter",
                          direction = "wide")

# Replace any NAs with 0
quarterly_wide[is.na(quarterly_wide)] <- 0

# Rename columns like revenue.1 → Q1
colnames(quarterly_wide) <- sub("revenue\\.1", "Q1", colnames(quarterly_wide))
colnames(quarterly_wide) <- sub("revenue\\.2", "Q2", colnames(quarterly_wide))
colnames(quarterly_wide) <- sub("revenue\\.3", "Q3", colnames(quarterly_wide))
colnames(quarterly_wide) <- sub("revenue\\.4", "Q4", colnames(quarterly_wide))

# Add total and average
quarter_cols <- c("Q1", "Q2", "Q3", "Q4")
quarterly_wide$Total <- rowSums(quarterly_wide[, quarter_cols])
quarterly_wide$Average <- round(quarterly_wide$Total / 4)

# Order by total and select top 5
top5_quarterly <- head(quarterly_wide[order(-quarterly_wide$Total), ], 5)

# Final column ordering
top5_quarterly <- top5_quarterly[, c("country", "Q1", "Q2", "Q3", "Q4", "Average")]

# Show the table
kbl(top5_quarterly,
    caption = sprintf("Cumulative and Average Quarterly Revenue (from %d to %d)", start_year, end_year),
    col.names = c("Country", "Q1", "Q2", "Q3", "Q4", "Average"),
    align = "lrrrrr") %>%
  kable_styling("striped", full_width = FALSE, position = "center")

```


#### 2.2 Customer Distribution

```{r echo = FALSE, message=FALSE, warning=FALSE}
# Aggregate customers by country and media_type
customers_by_unit <- dbGetQuery(con, "
  SELECT country, media_type, COUNT(DISTINCT customer_id) AS num_customers
  FROM dim_customer
  GROUP BY country, media_type;
")

# Reshape to wide (film, music) per country
customers_wide <- reshape(customers_by_unit,
                          idvar = "country",
                          timevar = "media_type",
                          direction = "wide")

# Replace NA with 0
customers_wide[is.na(customers_wide)] <- 0

# Rename columns to simpler names (remove prefix "num_customers.")
colnames(customers_wide) <- sub("num_customers\\.", "", colnames(customers_wide))

# Calculate total customers per country (sum film + music)
customers_wide$Total <- customers_wide$film + customers_wide$music

# Select top 5 countries by total customers
top5_customers <- head(customers_wide[order(-customers_wide$Total), ], 5)

kbl(top5_customers,
    caption = "Customers by Business Unit",
    col.names = c("Country", "Film", "Music", "Total"),
    align = "lrrr") %>%
  kable_styling("striped", full_width = FALSE, position = "center")


```



#### 2.3 Film vs Music Revenue

Sales fluctuate over time and the table below shows total revenue per month for the years for which we have data.
```{r echo = FALSE, message=FALSE, warning=FALSE}
# Query revenue by media_type and year
revenue_by_year_unit <- dbGetQuery(con, "
  SELECT 
    media_type,
    YEAR(date) AS year,
    ROUND(SUM(revenue), 0) AS revenue
  FROM fact_sales
  GROUP BY media_type, year
  ORDER BY media_type, year;
")

# Reshape wide: rows = media_type, columns = years
revenue_wide <- reshape(revenue_by_year_unit,
                        idvar = "media_type",
                        timevar = "year",
                        direction = "wide")

# Replace NAs with zeros
revenue_wide[is.na(revenue_wide)] <- 0

# Identify revenue columns
rev_cols <- grep("^revenue\\.", colnames(revenue_wide))

# Calculate total revenue per media_type across all years
revenue_wide$Total <- rowSums(revenue_wide[, rev_cols])

# Rename columns: strip "revenue." prefix
colnames(revenue_wide)[rev_cols] <- sub("revenue\\.", "", colnames(revenue_wide)[rev_cols])

# Capitalize media_type names for display
revenue_wide$media_type <- tools::toTitleCase(revenue_wide$media_type)

# Reorder columns: media_type first, then years in ascending order, then Total
year_cols <- sort(as.numeric(colnames(revenue_wide)[rev_cols]))
col_order <- c("media_type", as.character(year_cols), "Total")
revenue_wide <- revenue_wide[, col_order]

# Format numbers with commas
for (col in col_order[-1]) {
  revenue_wide[[col]] <- formatC(revenue_wide[[col]], format = "d", big.mark = ",")
}

# Show table with kableExtra
kbl(revenue_wide,
    caption = "Revenue by Year and Business Unit",
    col.names = c("Business Unit", as.character(year_cols), "Total"),
    align = c("l", rep("r", length(year_cols) + 1)),
    digits = 0) %>%
  kable_styling("striped", full_width = FALSE, position = "center")

```

The graph below illustrates the quarterly growth of the film and music business over the past years for which we have data.
```{r echo = FALSE, message=FALSE, warning=FALSE}

# 获取季度收入数据
quarterly_data <- dbGetQuery(con, "
  SELECT 
    YEAR(date) AS year,
    QUARTER(date) AS quarter,
    media_type,
    ROUND(SUM(revenue), 0) AS revenue
  FROM fact_sales
  GROUP BY year, quarter, media_type
  ORDER BY year, quarter, media_type;
")

# 创建 year_quarter 列
quarterly_data$year_quarter <- paste(quarterly_data$year, paste0("Q", quarterly_data$quarter), sep = " ")

# 创建一个排序用的辅助列
quarterly_data$year_quarter_order <- quarterly_data$year * 10 + quarterly_data$quarter

# 取出所有唯一的 year_quarter，并按顺序排序
year_quarter_levels <- unique(quarterly_data[order(quarterly_data$year_quarter_order), "year_quarter"])

# 转换为有序 factor
quarterly_data$year_quarter <- factor(quarterly_data$year_quarter, levels = year_quarter_levels, ordered = TRUE)

# 筛选 film 和 music 两种类型的数据
selected_data <- subset(quarterly_data, media_type %in% c("film", "music"))

# 作图
library(ggplot2)
library(scales)

ggplot(selected_data, aes(x = year_quarter, y = revenue, group = media_type, color = media_type)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(title = "Quarterly Revenue by Media Type",
       x = "Year - Quarter",
       y = "Revenue (USD)") +
  scale_y_continuous(labels = scales::comma_format()) +
  theme_minimal() +
  facet_wrap(~media_type, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")


ggplot(selected_data, aes(x = year_quarter, y = revenue, group = media_type, color = media_type)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(title = "Quarterly Revenue for Film and Music (Log Scale)",
       x = "Year - Quarter",
       y = "Revenue (USD, log scale)",
       color = "Media Type") +
  scale_y_log10(labels = scales::comma_format()) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))




```

In terms of units sold, the table below sheds light on this by Top 3 country and by business unit.

```{r echo = FALSE, message=FALSE, warning=FALSE}
# Step 1: Get the most recent 3 years
most_recent_years <- dbGetQuery(con, "
  SELECT DISTINCT YEAR(date) AS year
  FROM fact_sales
  ORDER BY YEAR(date) DESC
  LIMIT 3;
")

years <- sort(most_recent_years$year)
year_names <- as.character(years)
quarters <- paste0("Q", 1:4)

# Step 2: Query sales data
units_data <- dbGetQuery(con, sprintf("
  SELECT
    country,
    YEAR(date) AS year,
    QUARTER(date) AS quarter,
    SUM(units_sold) AS units_sold
  FROM fact_sales
  WHERE YEAR(date) IN (%d, %d, %d)
  GROUP BY country, YEAR(date), QUARTER(date)
  ORDER BY country, year, quarter;",
  years[1], years[2], years[3]
))

# Step 3: Calculate country totals and get top 3
country_totals <- aggregate(units_sold ~ country, data = units_data, sum)
top3_countries <- head(country_totals[order(-country_totals$units_sold), "country"], 3)

# Build final table (top 3 countries)
final_table <- data.frame()

for (country in top3_countries) {
  country_table <- matrix(0, nrow = 6, ncol = length(year_names) + 2)
  rownames(country_table) <- c(quarters, "Total", "Average")
  colnames(country_table) <- c(year_names, "Total", "Average")

  # Filter data for this country
  cdata <- units_data[units_data$country == country, ]

  # Fill in quarterly sales
  for (q in 1:4) {
    for (y in years) {
      val <- cdata$units_sold[cdata$quarter == q & cdata$year == y]
      if (length(val) == 1) {
        country_table[q, as.character(y)] <- val
      }
    }
  }

  # Calculate row totals and averages
  for (q in 1:4) {
    country_table[q, "Total"] <- sum(country_table[q, year_names])
    country_table[q, "Average"] <- round(mean(country_table[q, year_names]))
  }

  # Calculate column totals and averages
  for (y in year_names) {
    country_table["Total", y] <- sum(country_table[quarters, y])
    country_table["Average", y] <- round(mean(country_table[quarters, y]))
  }
  
  # Calculate grand totals
  country_table["Total", "Total"] <- sum(country_table[quarters, "Total"])
  country_table["Average", "Total"] <- round(mean(country_table[quarters, "Total"]))
  country_table["Total", "Average"] <- round(mean(country_table["Total", year_names]))
  country_table["Average", "Average"] <- round(mean(country_table[quarters, "Average"]))

  # Convert to data.frame
  df_country <- as.data.frame(country_table)
  df_country$Quarter <- rownames(country_table)
  
  # Reorder columns
  df_country <- df_country[, c("Quarter", year_names, "Total", "Average")]
  
  # Add country header row
  country_header <- data.frame(
    Quarter = country,
    matrix("", nrow = 1, ncol = ncol(df_country) - 1),
    stringsAsFactors = FALSE
  )
  names(country_header) <- names(df_country)
  
  # Combine
  final_table <- rbind(final_table, country_header, df_country)
}

# Clean up row names
rownames(final_table) <- NULL

# Step 5: Print table with proper formatting
tbl <- kbl(
  final_table,
  caption = "Number of units sold by country and quarter for past three years",
  col.names = c("", year_names, "Total", "Average"),
  align = c("l", rep("r", length(year_names) + 2)),
  format.args = list(big.mark = ","),
  escape = FALSE
) %>%
  kable_styling("striped", full_width = FALSE, position = "center") %>%
  row_spec(which(nchar(final_table$Quarter) > 2),  # Country rows
          bold = TRUE, 
          background = "#f2f2f2",
          extra_css = "border-bottom: 1px solid #ddd;") %>%
  column_spec(1, bold = TRUE) %>%
  collapse_rows(columns = 1, target = 1, valign = "top")

tbl
```
### 3. Summary and Recommendations
```{r echo = FALSE, message=FALSE, warning=FALSE}
# 1. Get min and max year of sales data
year_range <- dbGetQuery(con, "
  SELECT MIN(YEAR(date)) AS min_year, MAX(YEAR(date)) AS max_year
  FROM fact_sales;
")
min_year <- year_range$min_year
max_year <- year_range$max_year

# 2. Year with peak sales and total revenue for that year
peak_year_data <- dbGetQuery(con, "
  SELECT YEAR(date) AS year, ROUND(SUM(revenue), 0) AS total_revenue
  FROM fact_sales
  GROUP BY year
  ORDER BY total_revenue DESC
  LIMIT 1;
")
peak_year <- peak_year_data$year
peak_revenue <- formatC(peak_year_data$total_revenue, format = "d", big.mark = ",")

# 3. Total revenue by media_type
media_revenue <- dbGetQuery(con, "
  SELECT media_type, ROUND(SUM(revenue), 0) AS total_revenue
  FROM fact_sales
  GROUP BY media_type;
")
music_revenue_val <- media_revenue$total_revenue[media_revenue$media_type == "music"]
film_revenue_val <- media_revenue$total_revenue[media_revenue$media_type == "film"]

music_revenue <- formatC(music_revenue_val, format = "d", big.mark = ",")
film_revenue <- formatC(film_revenue_val, format = "d", big.mark = ",")

total_revenue_all <- sum(media_revenue$total_revenue)
music_pct <- round(music_revenue_val / total_revenue_all * 100, 1)
film_pct <- round(film_revenue_val / total_revenue_all * 100, 1)

# 4. Top 3 countries by customer count
top_countries <- dbGetQuery(con, "
  SELECT country, COUNT(DISTINCT customer_id) AS customer_count
  FROM dim_customer
  GROUP BY country
  ORDER BY customer_count DESC
  LIMIT 3;
")
top_countries_list <- paste(top_countries$country, collapse = ", ")

# Get total distinct customers
total_customers <- dbGetQuery(con, "SELECT COUNT(DISTINCT customer_id) AS total_customers FROM dim_customer;")$total_customers
total_customers_fmt <- formatC(total_customers, format = "d", big.mark = ",")

# Get total distinct countries customers come from
total_countries <- dbGetQuery(con, "SELECT COUNT(DISTINCT country) AS total_countries FROM dim_customer;")$total_countries


```

Based on the available data, the analysis of Media Distributors, Inc.’s film and music sales reveals a robust revenue generation pattern, with Canada and the United States emerging as the primary contributors to the company’s total sales. Between `r min_year` and `r max_year`, sales peaked in `r peak_year`, with total revenue reaching $`r peak_revenue` across both business units. Music and film sales contribute $`r music_revenue` and $`r film_revenue` respectively, accounting for approximately `r music_pct`% and `r film_pct`% of total revenue. The customer base is concentrated in `r top_countries_list`, which also serve as the top-performing markets.

Despite these strengths, Media Distributors, Inc. faces challenges related to its siloed information systems. With plans for acquisition, the company must present a cohesive and integrated business view to appeal to potential buyers. With nearly `r total_customers_fmt` customers across `r total_countries` countries, Media Distributors should invest in customer relationship management (CRM) tools to personalize outreach and foster loyalty. Segmenting customers by preferences (film or music) and spending habits can optimize targeting.



```{r disconnect, include=FALSE}
dbDisconnect(con)
```
