---
title: "Explore Data Warehouses"
author: Xujia Qin
semester: SummerI 2025
date: Jun 15th, 2025
output:
  pdf_document: 
    latex_engine: xelatex
    extra_dependencies: ["booktabs", "longtable", "array", "multirow", "xcolor", "wrapfig", "float", "colortbl"]
---

### Q1: Fact Tables, Star Schemas, and OLAP in Relational Databases

#### Fact Tables
- Store **numeric business metrics** (e.g., sales, revenue, quantities).  
- Each row represents a measurable event (like a transaction or shipment).  
- Use **foreign keys** to link to **dimension tables** (e.g., Date, Product, Customer).  

#### Star Schema
- A simple, denormalized design with:  
  - A **central fact table** (containing the measurable data).  
  - Connected directly to **dimension tables** (descriptive attributes).  
- Visually resembles a **star**—hence the name!  
- Optimized for fast querying and reporting.  

#### Can a Transactional Database be Used for OLAP?

- Transactional (OLTP) databases are optimized for speed and data integrity during inserts/updates—great for daily business operations. However, they are **not suitable** for OLAP due to:

  - Poor performance with large, complex queries.
  - Normalized schemas not optimized for aggregations with redundent data
  - Overwhelming in CPU workload, OOM, or I/O bottleneck issues.
- Therefore, **OLTP systems should not be used to house OLAP**. Instead, ETL implementation is good into a dedicated warehouse design.

#### Why Use a Relational DB + Star Schema vs. NoSQL?

- While **NoSQL** systems (like MongoDB, Cassandra) offer horizontal scalability and flexibility, **relational databases with star schemas** are still preferred for OLAP because:
  
  - Powerful **query optimization** and **indexing**
  - Strong **data integrity** and constraints with ACID compliance
  - Better performance for complex **joins and aggregations** with star schemas
  - Mature **BI tool** (e.g., Tableau, Power BI, Looker)

---

### Q2: Data Warehouse vs. Data Mart vs. Data Lake

| Feature          | Data Warehouse 🏢           | Data Mart 🏪               | Data Lake 🌊               |
|------------------|----------------------------|---------------------------|---------------------------|
| **Who uses it?** | Big bosses & analysts      | Specific Department (e.g.:Marketing/Sales teams)     | Data engineers|
| **Data Type**    | Clean, organized data      | Clean, focused data       | Raw, messy data           |
| **Speed**        | Pretty fast                | Super fast                | Slow (needs ETL cleaning)     |
| **Cost**         | High                     | Moderate                       | Low                        |

## What They Actually Do

- **Data Warehouse**: Like a company's main library - all central repository for integrated, the important, cleaned-up data in one place.
- **Data Mart**: Like a department's bookshelf - just a focused subset of a data warehouse intended for a specific team.
- **Data Lake**: Like a giant storage unit for raw, unstructured, or semi-structured data.

## Real-Life Example
In my experience working on a data engineering internship,
1. Dumped all the raw data into an **AWS S3 data lake** (the "storage unit"as data lake)
2. Cleaned it up and moved it to **Snowflake** (the "main library" as datawarehouse)
3. Gave the marketing team their own **data mart** (the "special bookshelf")

## Video Reference (https://www.youtube.com/watch?v=-bSkREem8dM)

---

### Q3: Designing a Fact Table for the Sakila Database

#### 🎯 Analytical Goal

I aim to analyze **rental revenue and activity** over time, broken down by staff, customer, and store. This will support monthly/quarterly revenue trends.
---

### ⭐ Star Schema Overview

We design a fact table named `fact_rental_revenue` and connect it with four dimension tables:

| Table              | Type       | Description                                |
|-------------------|------------|--------------------------------------------|
| `fact_rental_revenue` | Fact      | Metrics such as revenue and late days       |
| `dim_date`         | Dimension  | Rental date in various formats             |
| `dim_customer`     | Dimension  | Customer information                       |
| `dim_staff`        | Dimension  | Staff who processed the rental             |
| `dim_store`        | Dimension  | Store location                             |

---

### 🏗️ Fact Table: `fact_rental_revenue`

Each row corresponds to one rental transaction.

| Column Name         | Type    | Description                                 |
|---------------------|---------|---------------------------------------------|
| `rental_id`         | INT     | Primary key (from `rental` table)           |
| `date_id`           | INT     | Foreign key to `dim_date`                   |
| `customer_id`       | INT     | Foreign key to `dim_customer`               |
| `staff_id`          | INT     | Foreign key to `dim_staff`                  |
| `store_id`          | INT     | Foreign key to `dim_store`                  |
| `rental_amount`     | REAL    | Amount paid for the rental (from `payment`)|
| `late_return_days`  | INT     | Return delay in days (return - rental date) |

---

### 📅 Dimension: `dim_date`

| Column       | Description            |
|--------------|------------------------|
| `date_id`    | Surrogate primary key  |
| `date`       | Date (YYYY-MM-DD)      |
| `day`        | Day of month           |
| `month`      | Month number           |
| `year`       | Year                   |
| `weekday`    | Weekday name           |
| `is_weekend` | Boolean                |

---

### 🧑 Dimension: `dim_customer`

| Column        | Description             |
|---------------|-------------------------|
| `customer_id` | Customer ID (PK)        |
| `first_name`  |                         |
| `last_name`   |                         |
| `email`       |                         |
| `active`      | TRUE/FALSE              |

---

### 🧍 Dimension: `dim_staff`

| Column       | Description              |
|--------------|--------------------------|
| `staff_id`   | Staff ID (PK)            |
| `first_name` |                          |
| `last_name`  |                          |
| `store_id`   | Foreign key to store     |

---

### 🏬 Dimension: `dim_store`

| Column       | Description             |
|--------------|-------------------------|
| `store_id`   | Store ID (PK)           |
| `address_id` | To join with location   |
| `manager_id` |                         |

---

### Q4: Create and Populate OLAP Fact Table in `SakilaOLAP.db`

In this step,
1. Create a new OLAP database: `SakilaOLAP.db`.
2. Design and create a fact table: `fact_rental_revenue`.
3. Extract relevant rental facts from the original `sakila.db`.
4. Populate the fact table in the new OLAP database.
---

---

### 1. Connect to Databases
```{r setup, include=FALSE}
# Load libraries
library(DBI)
library(RSQLite)
library(knitr)
```

```{r echo = FALSE, message=FALSE, warning=FALSE}
# Connect to original transactional database
sakila_conn <- dbConnect(SQLite(), "sakila.db")

# Create and connect to new OLAP database
olap_conn <- dbConnect(SQLite(), "SakilaOLAP.db")
print("Database Connected!")
```

---

### 2. Create Fact Table in OLAP Database

```{r echo = FALSE, message=FALSE, warning=FALSE}
# Drop table if it exists
invisible(dbExecute(olap_conn, "DROP TABLE IF EXISTS fact_rental_revenue"))

# Create a new fact table to store aggregated rental revenue information
create_fact_table_sql <- "
  CREATE TABLE fact_rental_revenue (
    rental_id INTEGER PRIMARY KEY,
    date_id TEXT,
    customer_id INTEGER,
    staff_id INTEGER,
    store_id INTEGER,
    rental_amount REAL,
    late_return_days INTEGER
  );
"

invisible(dbExecute(olap_conn, create_fact_table_sql))
print("Fact table is created!")
```

---

### 3. Extract and Transform Facts from `sakila.db`

I'll join the following tables:

- `rental` for rental info
- `payment` for rental revenue
- `inventory` for store ID
- `staff` and `customer` for foreign keys

```{r echo = FALSE, message=FALSE, warning=FALSE}
extract_query <- "
  SELECT 
    r.rental_id,
    DATE(r.rental_date) AS date_id,
    r.customer_id,
    r.staff_id,
    i.store_id,
    IFNULL(SUM(p.amount), 0) AS rental_amount,
    CAST((JULIANDAY(r.return_date) - JULIANDAY(r.rental_date)) AS INTEGER) AS late_return_days
  FROM rental r
  JOIN inventory i ON r.inventory_id = i.inventory_id
  LEFT JOIN payment p ON r.rental_id = p.rental_id
  GROUP BY r.rental_id
"

rental_facts <- dbGetQuery(sakila_conn, extract_query)
head(rental_facts)
```

---

### 4. Load Data into OLAP Fact Table

```{r echo = FALSE, message=FALSE, warning=FALSE}
# Insert extracted data into OLAP database
dbWriteTable(olap_conn, "fact_rental_revenue", rental_facts, append = TRUE, row.names = FALSE)

# Check that data was inserted
dbGetQuery(olap_conn, "SELECT COUNT(*) AS row_count FROM fact_rental_revenue")
```

---

### 5. Preview Final Fact Table

```{r echo = FALSE, message=FALSE, warning=FALSE}
# Display sample rows from OLAP fact table
dbGetQuery(olap_conn, "SELECT * FROM fact_rental_revenue LIMIT 10")
```

---

### Q5: Rental Analytics — Average Rental Amount Per Month (year 2005)


```{r echo = FALSE, message=FALSE, warning=FALSE}
query <- "
  SELECT 
    SUBSTR(date_id, 1, 7) AS year_month,
    ROUND(AVG(rental_amount), 2) AS avg_rental_amount
  FROM fact_rental_revenue
  WHERE date_id LIKE '2005-%'
  GROUP BY year_month
  ORDER BY year_month
"

monthly_avg <- dbGetQuery(olap_conn, query)
print(monthly_avg)
```
### formatted table
```{r echo = FALSE, message=FALSE, warning=FALSE}
kable(monthly_avg, col.names = c("Year-Month", "Average Rental Amount ($)"))
```


### Disconnect from Both Databases

```{r echo = FALSE, message=FALSE, warning=FALSE}
dbDisconnect(sakila_conn)
dbDisconnect(olap_conn)
print("Database is disconnected!")
```


