---
title: "R Notebook"
output: html_notebook
---
## MYSQL on Aiven
```{r}
# 1. install lib
install.packages("RMySQL")
install.packages("DBI")
```

```{r}
# 2. load lib
library(DBI)
library(RMySQL)


# 2. settings
db_user <- 'avnadmin'
db_password <- 'AVNS_myNMlu7wBu69NOuFkU1'
db_name <- 'defaultdb'
db_host <- 'mysql-1c2ba4a8-northeastern-cee5.k.aivencloud.com'
db_port <- 19461

# 3. read data from db
mydb <- dbConnect(MySQL(), user = db_user, password = db_password, 
                  dbname = db_name, host = db_host, port = db_port)


# Create a simple test table
dbExecute(mydb, "CREATE TABLE test_table (id INT PRIMARY KEY, name VARCHAR(50));")

# Insert some data
dbExecute(mydb, "INSERT INTO test_table VALUES (1, 'Alice'), (2, 'Bob');")

# Query the data back
result <- dbGetQuery(mydb, "SELECT * FROM test_table;")
print(result)

# Clean up (optional)
dbExecute(mydb, "DROP TABLE test_table;")




```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

